---
title: "Seurat workflow male mesonephros"
output:
  html_document:
    keep_md: true
    smart: false
    toc: true
    toc_float: true
    theme: united
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache.lazy = FALSE,
  tidy = TRUE
)
```

# Male mesonephros 
* Workflow with output on both batch corrected and non batch corrected 
* Non corrected output is retaining useful information compared to the corrected 

## Libraries
```{r}
suppressMessages(library(plotly))
suppressMessages(library(Seurat))
suppressMessages(library(dplyr))
suppressMessages(library(Matrix))
suppressMessages(library(topGO))
suppressMessages(library(org.Hs.eg.db))
suppressMessages(library(gplots))
suppressMessages(library(genefilter))
suppressMessages(library(future))
suppressMessages(library(batchelor))
suppressMessages(library(SeuratWrappers))
suppressMessages(library(gprofiler2))
suppressMessages(library(plotly))
suppressMessages(library(gridExtra))

## We can use multiple cores for some functions, see: https://satijalab.org/seurat/v3.2/future_vignette.html
plan("multiprocess", workers = 4)
plan()

outputDir = getwd()

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

## Load Seurat object
* Male only embryos separated in this object
```{r}
object <- readRDS("/path/to/mesonephros_male.rds")
```

## Add age info
* Weeks post fertilization
  * TK1 <- wpf_07
  * TN1 <- wpf_11
  * TR5 <- wpf_17
```{r}
subjects <- object@meta.data$cleaned_subject_ids
subjects["TK1"] <- "wpf_07"
subjects["TN1"] <- "wpf_11"
subjects["TR5"] <- "wpf_17"
names(subjects) <- colnames(object)
object[["age"]] <- subjects
```

## QC
* Remove cells expressing less than 750 genes, more than 7000 genes and have more than 50000 UMIs
* Remove cells with more than 10% MT 
* Calculate the percentage of cell cycle genes (same function as for MT genes)
```{r}
Idents(object) <- "mesonephros"
object[["percent.mt"]] <- PercentageFeatureSet(object, 
                                                      pattern = "^MT-")
cc_genes_in_dataset <- c(s.genes, g2m.genes)[c(s.genes, g2m.genes) %in% rownames(object)]
object[["percent.cc_genes"]] <- PercentageFeatureSet(object, 
                                                    features = cc_genes_in_dataset)
p1 <- VlnPlot(object, 
             features = "nFeature_RNA", 
             ncol = 1,
             pt.size = 0)
p1 <- p1 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
p2 <- VlnPlot(object, 
             features = "nCount_RNA", 
             ncol = 1,
             pt.size = 0)
p2 <- p2 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
p3 <- VlnPlot(object, 
              features = "percent.mt", 
              ncol = 1,
              pt.size = 0)
p3 <- p3 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
grid.arrange(p1, p2, p3, ncol=3)
object <- subset(object, 
                        subset = nFeature_RNA > 750 & nFeature_RNA <7000 & percent.mt < 10 & nCount_RNA < 50000)
p1 <- VlnPlot(object, 
             features = "nFeature_RNA", 
             ncol = 1,
             pt.size = 0)
p1 <- p1 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
p2 <- VlnPlot(object, 
             features = "nCount_RNA", 
             ncol = 1,
             pt.size = 0)
p2 <- p2 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
p3 <- VlnPlot(object, 
              features = "percent.mt", 
              ncol = 1,
              pt.size = 0)
p3 <- p3 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
grid.arrange(p1, p2, p3, ncol=3)
object
```

## Calculate cell cycle scores
* Sum the expression of cell cycle genes
```{r}
object <- CellCycleScoring(object, 
                                s.features = s.genes, 
                                g2m.features = g2m.genes, 
                                set.ident = TRUE)
```

## Dim reduction
```{r}
object <- RunPCA(object, 
                        verbose = FALSE)
object <- FindNeighbors(object, 
                               dims = c(1:15))
object <- FindClusters(object,
                              resolution = 0.5)
object <- RunUMAP(object, 
                         dims = c(1:15))
DimPlot(object, 
             reduction = "umap")
ggsave("umap_clusters.pdf",
       width = 10,
       height = 7)
DimPlot(object, 
             reduction = "umap",
             group.by = "cleaned_subject_ids")
ggsave("umap_subjects.pdf",
       width = 10,
       height = 7)
DimPlot(object, 
             reduction = "umap",
             group.by = "age")
ggsave("umap_age.pdf",
       width = 10,
       height = 7)
DimPlot(object, 
             reduction = "umap",
             group.by = "Phase")
ggsave("umap_phase.pdf",
       width = 10,
       height = 7)
FeaturePlot(object,
            reduction = "umap",
            feature= "percent.cc_genes")
FeaturePlot(object,
            reduction = "umap",
            feature= "percent.dissoc")
object[["Clusters_before_mnn"]] <- Idents(object)
saveRDS(object, 
        file = "mesonephros_male_no_mnn.rds")
```
`
### Session Info
```{r}
sessionInfo()
```
