---
title: "Standard Workflow J. Guo, own male gonads and male mesonephros"
output:
  html_document:
    keep_md: true
    smart: false
    toc: true
    toc_float: true
    theme: united
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache.lazy = FALSE,
  tidy = TRUE
)
```

# J. Guo testis dataset
* 10.1016/j.stem.2020.12.004

## Libraries
```{r}
suppressMessages(library(plotly))
suppressMessages(library(Seurat))
suppressMessages(library(dplyr))
suppressMessages(library(Matrix))
suppressMessages(library(topGO))
suppressMessages(library(org.Hs.eg.db))
suppressMessages(library(gplots))
suppressMessages(library(genefilter))
suppressMessages(library(future))
suppressMessages(library(batchelor))
suppressMessages(library(SeuratWrappers))
suppressMessages(library(gprofiler2))
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))

## We can use multiple cores for some functions, see: https://satijalab.org/seurat/v3.2/future_vignette.html
plan("multiprocess", workers = 2)
plan()
outputDir = getwd()

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

## Load Seurat objects
* Guo + in house male gonads. For details refer to: https://github.com/johnmous/single_cell-embryo_gonads/blob/main/male_gonads_plus_guo.Rmd
* Male mesonephros. For details refer to: https://github.com/johnmous/single_cell-embryo_gonads/blob/main/male_mesonephros.Rmd
```{r}
guo_in_house_gonads <- readRDS('/path/to/combined_mnn_pcs15_res_0.4.rds')
Idents(guo_in_house_gonads) <- 'Gonads'
male_mesonephros <- readRDS('/path/to/mesonephros_male_no_mnn.rds')
Idents(male_mesonephros) <- 'Mesonephros'
male_mesonephros[['origin']] <- 'mesonephros'
male_mesonephros[['origin_trim']] <- 'mesonephros'
```

## Rename age of mesonephros for consistency
```{r}
age_wpf <- male_mesonephros[["age"]]
age_wpf[age_wpf == "TK1"] <- "wpf_07"
age_wpf[age_wpf == "TN1"] <- "wpf_11"
age_wpf[age_wpf == "TR5"] <- "wpf_17"
male_mesonephros[['age']] <- age_wpf
table(male_mesonephros[['age']] )
```

## Merge Seurat objects
* Annotate by age (Weeks Post Fertilization)
```{r}
gonads_mesonephros <- merge(x = guo_in_house_gonads, 
                            y = male_mesonephros, 
                            add.cell.ids = c("gonads", "mesonephros"))
table(gonads_mesonephros[['orig.ident']])
table(gonads_mesonephros[['age']])
```

## QC
* Visualize QC previously applied
```{r}
p1 <- VlnPlot(gonads_mesonephros, 
             features = "nFeature_RNA", 
             ncol = 1,
             pt.size = 0)
p1 <- p1 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
p2 <- VlnPlot(gonads_mesonephros, 
             features = "nCount_RNA", 
             ncol = 1,
             pt.size = 0)
p2 <- p2 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
p3 <- VlnPlot(gonads_mesonephros, 
              features = "percent.mt", 
              ncol = 1,
              pt.size = 0)
p3 <- p3 + geom_jitter(size = 0.1, alpha = 0.2) + ylim(0, NA) + theme(legend.position = "none")
grid.arrange(p1, p2, p3, ncol=3)
```

## Normalize and scale data
```{r}
gonads_mesonephros <- NormalizeData(gonads_mesonephros, 
                                    normalization.method = "LogNormalize", 
                                    scale.factor = 50000)
gonads_mesonephros <- FindVariableFeatures(gonads_mesonephros, 
                                           selection.method = "vst")
gonads_mesonephros <- ScaleData(gonads_mesonephros) 
```

## Run PCA, UMAP, clustering functions
```{r}
gonads_mesonephros <- RunPCA(gonads_mesonephros, 
                             verbose = FALSE)
ElbowPlot(gonads_mesonephros)
gonads_mesonephros <- FindNeighbors(gonads_mesonephros, 
                                    dims = c(1:15))
gonads_mesonephros <- FindClusters(gonads_mesonephros,
                                   resolution = 0.4)
gonads_mesonephros <- RunUMAP(gonads_mesonephros, 
                              dims = c(1:15))
DimPlot(gonads_mesonephros, 
        reduction = "umap")
ggsave("umap_clusters.pdf",
       width = 10,
       height = 7)
DimPlot(gonads_mesonephros, 
        reduction = "umap",
        group.by = "age")
ggsave("umap_clusters_age.pdf",
       width = 10,
       height = 7)
DimPlot(gonads_mesonephros, 
        reduction = "umap",
        group.by = "origin")
ggsave("umap_clusters_origin.pdf",
       width = 10,
       height = 7)
saveRDS(object = gonads_mesonephros, 
        file = "gonads_mesonephros_no_mnn.rds")
```

## Batch corrrect MNN between in house and guo
* 
```{r}
gonads_mesonephros <- RunFastMNN(object.list = SplitObject(gonads_mesonephros, split.by = "origin_trim"))
```

## Redo with mnn correction
* Use 30 PCs for the cluster and umap calculations, as in the paper???
```{r}
gonads_mesonephros <- FindNeighbors(gonads_mesonephros, 
                                    reduction = "mnn",
                                    dims = c(1:15))
gonads_mesonephros <- FindClusters(gonads_mesonephros,
                                   resolution = 0.4)
gonads_mesonephros <- RunUMAP(gonads_mesonephros,
                              reduction = "mnn",
                              dims = c(1:15))
DimPlot(gonads_mesonephros, 
        reduction = "umap")
ggsave("umap_clusters_mnn.pdf",
       width = 10,
       height = 7)
DimPlot(gonads_mesonephros, 
        reduction = "umap",
        group.by = "age")
ggsave("umap_clusters_age_mnn.pdf",
       width = 10,
       height = 7)
DimPlot(gonads_mesonephros, 
        reduction = "umap",
        group.by = "origin")
ggsave("umap_clusters_origin_mnn.pdf",
       width = 10,
       height = 7)
saveRDS(object = gonads_mesonephros, 
        file = "gonads_mesonephros_mnn.rds")
```

## Find markers
```{r}
markers <- FindAllMarkers(object = object,
                          only.pos = TRUE,
                          min.pct = 0.25)
filtered_markers <- markers %>% filter(pct.1>0.6 & p_val_adj <0.05 )
allMarkersPath = paste0(outputDir, "/gonads_mesonephros_mnn_markers_filtered.tsv")
write.table(x = filtered_markers,
            file = allMarkersPath,
            sep = "\t",
            quote = FALSE,
            row.names = FALSE)
```


### Session Info
```{r}
sessionInfo()
```
