---
title: "Standard Workflow J. Guo dataset plus in house"
output:
  html_document:
    keep_md: true
    smart: false
    toc: true
    toc_float: true
    theme: united
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache.lazy = FALSE,
  tidy = TRUE
)
```

# J. Guo testis dataset
* 10.1016/j.stem.2020.12.004

## Libraries
```{r}
suppressMessages(library(plotly))
suppressMessages(library(Seurat))
suppressMessages(library(dplyr))
suppressMessages(library(Matrix))
suppressMessages(library(topGO))
suppressMessages(library(org.Hs.eg.db))
suppressMessages(library(gplots))
suppressMessages(library(genefilter))
suppressMessages(library(future))
suppressMessages(library(batchelor))
suppressMessages(library(SeuratWrappers))
suppressMessages(library(gprofiler2))

## We can use multiple cores for some functions, see: https://satijalab.org/seurat/v3.2/future_vignette.html
plan("multiprocess", workers = 2)
plan()
options(future.globals.maxSize= 1991289600)
outputDir = getwd()

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

## Load 
* There are 6 separate ages
```{r}
week_6_data <- Read10X(data.dir = "/path/to/j_guo_male_testis/male_week_6")
week_7_data <- Read10X(data.dir = "/path/to/j_guo_male_testis/male_week_7")
week_8_data <- Read10X(data.dir = "/path/to/j_guo_male_testis/male_week_8")
week_12_data <- Read10X(data.dir = "/path/to/j_guo_male_testis/male_week_12")
week_15_data <- Read10X(data.dir = "/path/to/j_guo_male_testis/male_week_15")
week_16_data <- Read10X(data.dir = "/path/to/j_guo_male_testis/male_week_16")
```

## Create seurat objects
```{r}
week_6 <- CreateSeuratObject(counts = week_6_data, project = "wpf_06")
week_7 <- CreateSeuratObject(counts = week_7_data, project = "wpf_07")
week_8 <- CreateSeuratObject(counts = week_8_data, project = "wpf_08")
week_12 <- CreateSeuratObject(counts = week_12_data, project = "wpf_12")
week_15 <- CreateSeuratObject(counts = week_15_data, project = "wpf_15")
week_16 <- CreateSeuratObject(counts = week_16_data, project = "wpf_16")
```



## Merge Seurat objects
* Annotate by age (Weeks Post Fertilization)
```{r}
guo_testis <- merge(x = week_6, 
                    y = c(week_7, week_8, week_12, week_15, week_16), 
                    add.cell.ids = c("wpf_06", "wpf_07", "wpf_08", "wpf_12", "wpf_15", "wpf_16"))
table(guo_testis$orig.ident)
```

## QC
* Remove cells expressing less than 750 genes and more than 5000
* Remove cells with more than 10% MT 
* Remove cells with more than 25000 UMIs
```{r}
guo_testis[["percent.mt"]] <- PercentageFeatureSet(guo_testis, 
                                                   pattern = "^MT-")
cc_genes_in_dataset <- c(s.genes, g2m.genes)[c(s.genes, g2m.genes) %in% rownames(guo_testis)]
guo_testis[["percent.cc_genes"]] <- PercentageFeatureSet(guo_testis, 
                                                    features = cc_genes_in_dataset)
VlnPlot(guo_testis, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 3,
        pt.size = 0.1)
guo_testis <- subset(guo_testis, 
                        subset = nFeature_RNA > 750 & nFeature_RNA < 7000 & percent.mt < 10 & nCount_RNA < 50000)
VlnPlot(guo_testis, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 4,
        pt.size = 0.3)
table(guo_testis$orig.ident)
guo_testis[['age']] <- Idents(guo_testis)
```

## Dissociation genes on normalizeds data
* Calculate the expression of dissociation genes as a fraction of total expression (for details see 10.1038/nmeth.4437)
* We calculate it on normalized counts
```{r}
guo_testis <- NormalizeData(guo_testis, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 50000)

genesChrom <- c("Actg1__chr11","Ankrd1__chr19","Arid5a__chr1","Atf3__chr1","Atf4__chr15","Bag3__chr7","Bhlhe40__chr6",
"Brd2__chr17","Btg1__chr10","Btg2__chr1","Ccnl1__chr3","Ccrn4l__chr3","Cebpb__chr2","Cebpd__chr16",
"Cebpg__chr7","Csrnp1__chr9","Cxcl1__chr5","Cyr61__chr3","Dcn__chr10","Ddx3x__chrX","Ddx5__chr11",
"Des__chr1","Dnaja1__chr4","Dnajb1__chr8","Dnajb4__chr3","Dusp1__chr17","Dusp8__chr7",
"Egr1__chr18","Egr2__chr10","Eif1__chr11","Eif5__chr12","Erf__chr7","Errfi1__chr4","Fam132b__chr1",
"Fos__chr12","Fosb__chr7","Fosl2__chr5","Gadd45a__chr6","Gcc1__chr6","Gem__chr4","H3f3b__chr11",
"Hipk3__chr2","Hsp90aa1__chr12","Hsp90ab1__chr17","Hspa1a__chr17","Hspa1b__chr17","Hspa5__chr2",
"Hspa8__chr9","Hspb1__chr5","Hsph1__chr5","Id3__chr4","Idi1__chr13","Ier2__chr8","Ier3__chr17",
"Ifrd1__chr12","Il6__chr5","Irf1__chr11","Irf8__chr8","Itpkc__chr7","Jun__chr4","Junb__chr8",
"Jund__chr8","Klf2__chr8","Klf4__chr4","Klf6__chr13","Klf9__chr19","Litaf__chr16","Lmna__chr3",
"Maff__chr15","Mafk__chr5","Mcl1__chr3","Midn__chr10","Mir22hg__chr11","Mt1__chr8","Mt2__chr8",
"Myadm__chr7","Myc__chr15","Myd88__chr9","Nckap5l__chr15","Ncoa7__chr10","Nfkbia__chr12","Nfkbiz__chr16",
"Nop58__chr1","Nppc__chr1","Nr4a1__chr15","Odc1__chr12","Osgin1__chr8","Oxnad1__chr14","Pcf11__chr7",
"Pde4b__chr4","Per1__chr11","Phlda1__chr10","Pnp__chr14","Pnrc1__chr4","Ppp1cc__chr5","Ppp1r15a__chr7",
"Pxdc1__chr13","Rap1b__chr10","Rassf1__chr9","Rhob__chr12","Rhoh__chr5","Ripk1__chr13","Sat1__chrX",
"Sbno2__chr10","Sdc4__chr2","Serpine1__chr5","Skil__chr3","Slc10a6__chr5","Slc38a2__chr15",
"Slc41a1__chr1","Socs3__chr11","Sqstm1__chr11","Srf__chr17","Srsf5__chr12","Srsf7__chr17",
"Stat3__chr11","Tagln2__chr1","Tiparp__chr3","Tnfaip3__chr10","Tnfaip6__chr2","Tpm3__chr3",
"Tppp3__chr8","Tra2a__chr6","Tra2b__chr16","Trib1__chr15","Tubb4b__chr2","Tubb6__chr18",
"Ubc__chr5","Usp2__chr9","Wac__chr18","Zc3h12a__chr4","Zfand5__chr19","Zfp36__chr7","Zfp36l1__chr12",
"Zfp36l2__chr17","Zyx__chr6","Gadd45g__chr13","Hspe1__chr1","Ier5__chr1","Kcne4__chr1")

genes <- sapply(genesChrom, function(x){
  toupper( strsplit(x, "__")[[1]][1])
})

Data <- as.data.frame(as.matrix(GetAssayData(object = guo_testis)))
cat("All genes:\n")
print(unname(genes))
write.table(genes, paste0(outputDir, "/mouseDissocGenes.tsv"), sep ="\t", quote=FALSE, row.names = FALSE)

## Remove mouse only genes and put the corresponding human
genes <- genes[!genes %in% c("CCRN4L", "MT1", "MT2")]
genes <- c(genes, "NOCT", "MT1A", "MT2A")
cat("Genes from mouse we miss in human:\n")
unname(genes[!genes %in% row.names(Data)])

## Calculate the percentage of UMIs maping on dissociation genes
totalSum <- Matrix::colSums(GetAssayData(object = guo_testis))
selection <- Data[genes, ]
selection[is.na(selection)] <- 0
dissociationSums <- colSums(selection)  
countSums <- merge(totalSum, dissociationSums, by="row.names", all=TRUE, sort= FALSE)
rownames(countSums) <- countSums$Row.names
countSums <- countSums[-1]
colnames(countSums) <- c("totalCount", "dissociationCounts")
countSums$percentage <- countSums$dissociationCounts/countSums$totalCount
## Save in meta.data of object
guo_testis[["percent.dissoc"]] <- countSums$percentage

## Draw histogram for all samples
percentages <- guo_testis$percent.dissoc
hist(percentages, 
     breaks = 100, 
     col = "lightgrey", 
     main = paste("Expression dissociation-affected genes"), 
xlab = "Ratio of dissociation-affected genes to total gene count", ylab = "Number of cells", xlim = c(0, 0.20))

## Remove object to free up mem
rm(Data)
```

## Keep cells with dissociation percentages below the threshold of 6%
```{r}
guo_testis <- subset(x = guo_testis, 
                        subset = percent.dissoc < 0.06)
dim(guo_testis)
table(guo_testis$orig.ident)
```

## Combine with in house male gonads
* R object created from: https://github.com/johnmous/single_cell-embryo_gonads/blob/main/male_gonads.Rmd
```{r}
in_house_male_gonads <- readRDS("/path/to/male_gonads_no_mnn_15PCs.rds")
combined <- merge(x = guo_testis, y = in_house_male_gonads, add.cell.ids = c("guo", "in_house"))
```

## Redo analysis
```{r}
combined <- NormalizeData(combined, 
                            normalization.method = "LogNormalize", 
                            scale.factor = 50000)
combined <- FindVariableFeatures(combined, 
                                  selection.method = "vst")
combined <- ScaleData(combined) 
```

## Origin of data
```{r}
age <- combined[['age']]
age$origin <- age$age
age$origin[age$age %in% c('wpf_18', 'wpf_09a', 'wpf_09b')] <- "in_house"
age$origin[!(age$age %in% c('wpf_18', 'wpf_09a', 'wpf_09b'))] <- "guo"
combined[["origin"]] <- age$origin
```

## Origin of data, keep trimesters separate
* Keep the trimeters of own dataset separate
```{r}
age <- combined[['age']]
age$origin_trim <- age$age
age$origin_trim[!(age$age %in% c('wpf_18', 'wpf_09a', 'wpf_09b'))] <- "guo"
age$origin_trim[(age$age %in% c('wpf_18'))] <- "2nd"
age$origin_trim[(age$age %in% c('wpf_09a', 'wpf_09b'))] <- "1st"
combined[["origin_trim"]] <- age$origin_trim
```

## Use 30 PCs in the non-merged dataset
* Use 30 PCs for the cluster and umap calculations, as in the paper
```{r}
combined <- RunPCA(combined, 
                    verbose = FALSE)
ElbowPlot(combined)
combined <- FindNeighbors(combined, 
                               dims = c(1:15))
combined <- FindClusters(combined,
                              resolution = 0.4)
combined <- RunUMAP(combined, 
                         dims = c(1:15))
DimPlot(combined, 
        reduction = "umap")
ggsave("umap_clusters.pdf",
       width = 10,
       height = 7)
DimPlot(combined, 
        reduction = "umap",
        group.by = "age")
ggsave("umap_clusters_age.pdf",
       width = 10,
       height = 7)
DimPlot(combined, 
        reduction = "umap",
        group.by = "origin")
ggsave("umap_clusters_origin_mnn.pdf",
       width = 10,
       height = 7)
saveRDS(object = combined, 
        file = "combined_no_mnn.rds")
```

## Batch corrrect MNN between in house and guo
```{r}
combined <- RunFastMNN(object.list = SplitObject(combined, split.by = "origin_trim"))
```

## Redo with mnn correction
* Use 30 PCs for the cluster and umap calculations, as in the paper???
```{r}
combined <- FindNeighbors(combined, 
                          reduction = "mnn",
                          dims = c(1:15))
combined <- FindClusters(combined,
                              resolution = 0.4)
combined <- RunUMAP(combined,
                    reduction = "mnn",
                    dims = c(1:15))
DimPlot(combined, 
        reduction = "umap")
ggsave("umap_clusters_mnn.pdf",
       width = 10,
       height = 7)
DimPlot(combined, 
        reduction = "umap",
        group.by = "age")
ggsave("umap_clusters_age_mnn.pdf",
       width = 10,
       height = 7)
DimPlot(combined, 
        reduction = "umap",
        group.by = "origin")
ggsave("umap_clusters_origin_mnn.pdf",
       width = 10,
       height = 7)
saveRDS(object = combined, 
        file = "combined_mnn_pcs15_res_0.4.rds")
```


### Session Info
```{r}
sessionInfo()
```
